---
title: "Exploratory Data Analysis"
date: "07/10/2022"
author: "Jenny Yang"
output: 
  html_document:
  toc: true
  toc_float: true
---

## Introduction and set-up

In this section of the tutorial, we'll guide you through the process of Exploratory Data Analysis (EDA) using the Kiva dataset and prompt introduced in the **Getting Started** section. The purpose of EDA is to perform initial investigation on the data, look for interesting trends, relationships, and anomalies through the use of descriptive statistics and data visualization. The insights generated through this stage will motivate and inspire the subsequent analysis, modelling, and statistical tests performed at the next phase in your DataFest journey.

### Loading packages

Since you've already been through the **Getting Started** tutorial, your R environment is all set up! With the prompt in mind, let's take a look at the data. But first, let's load some useful packages. In this tutorial, we will mainly explore and visualize the data using `tidyverse`, a collection of incredibly useful packages for doing Data Science in $\texttt{R}$. Since countries in the dataset are recorded as country codes, the package `countrycode` will help to translate country code into country names. `colorblindr` and `viridis` are packages for colors in ggplot2.

```{r load-packages, message=FALSE, warning=FALSE, results=FALSE}
library(tidyverse)
library(countrycode)
library(colorblindr)
library(viridis)
```

### Overview of the Data

Recall the prompt: **Help understand what motivates people to lend money to developing-nation entrepreneurs and what factors are associated with paying these loans.** There are two parts of the prompt that we need to address, **motivation for lending** and **factors associated with repayment**.

From the file names and data dictionary, `lender_loans.csv`, `lenders.csv`, and `loans.csv` seem to be three most relevant files. We can take a quick peek at the data before diving more deeply in.

```{r read data, message = FALSE, warning = FALSE, results=FALSE}
lender_loans <- read_csv("data/lender_loans.csv")
lenders <- read_csv("data/lenders.csv")
loans <- read_csv("data/loans.csv")
str(lender_loans)
str(lenders)
str(loans)
```

Using the `str` function, we can see that in `lender_loans`, we have 125,580 observations of 51 variables; in `lenders`, we have 15,000 observations of 18 variables; and in `loans`, we have 97,183 observations of 75 variables. We can also see the variable names, from which we can narrow down our scope to the variables that are relevant to the prompt for further exploration and analysis.

To investigate lenders' motivations, we can focus on `lender_loans` and `lenders`. Possible helpful variables that associate with each lender can include `whereabouts` (introductory description about lenders), `country_code`(lenders' country), `occupation` (lenders' occupation), as well as the time that each lender has been a member of Kiva (variables start with `member_since`). `loan_because` will also be revealing since the lenders are directly stating their reasons for lending. Yet, it is a non-numerical data. So, it requires different analytical approaches that will be covered later in the **Text Analysis** section.

Similarly, variables that will contribute to the repayment can include `loan_amount`, `sector`, `use`, `location.code` (borrowers' country). A larger `loan_amount` may be a larger investment that is more risky. Certain sectors may be more profitable than others, similar with `use`, which is the detailed use of the money that borrowers present in their request. The macroeconomic environment of a location can be exceptionally critical to the success of borrowers.

There are also some variables that may be relevant to both parts of the prompt, such as the time when the loan request was posted and the loan amount. For instance, in 2008, the world economy underwent a global financial crisis. We might expect that this would have simultaneously increased the demand and decreased the supply of loans. Additionally, we might expect increased defaults as borrowers struggle to repay existing loans. To explore these hypotheses and more, we begin our analysis.

## EDA

We begin by visualizing the distribution of potentially important variables .

### -- LOOKING AT THE BIG PICTURE --

#### Loan Amount

Let's start simple. To get a general sense of the loan amount, we can plot a histogram with the `loan_amount` on the x-axis and the count on the y-axis. Histograms is one of the most basic and direct visualization graphs for continuous variables. We can also color the bars to see the proportion of loans in different statuses to add another layer of information.

```{r loan_amount dist, message = FALSE, warning = FALSE, fig.width = 8}
lender_loans$status <- gsub("in_repayment", "in repayment", lender_loans$status)
lender_loans$status <- sub("(.)", "\\U\\1",lender_loans$status,perl=TRUE)
ggplot(lender_loans, aes(x = loan_amount, fill = status, width = 0.1)) +
  geom_bar(aes(y=(..count..)/sum(..count..)), stat = "bin", color = "black") +
  xlab("Loan amount") +
  ylab("Percentage") +
  scale_fill_OkabeIto()+
  scale_fill_discrete(name = "Status")+
  ggtitle("Distribution of loan amount and the status of each loan amount")
```

From the graph above we can see that the loan amount has a unimodal, skewed right distribution. Over 50% of loans is concentrated below \$1000, while there are some large loans scattered on the right side of the graph. Through the color we can also tell that more than half of the loans are already paid, while a good proportion is still in repayment.

A summary table may be more direct to visualize the proportion of each status:

```{r table}

loans$status <- gsub("in_repayment", "in repayment", loans$status)
loans$status <- sub("(.)", "\\U\\1", loans$status, perl=TRUE)

status <- loans %>%
  drop_na(status) %>% # drop missing values by health variable
  group_by(status) %>% # specify categorical variable 
  rename("Status" = "status") %>% 
  summarize(Frequency = n()) %>% # return counts / frequencies
  arrange(desc(Frequency)) %>% 
  mutate(Percent = paste0(round(Frequency / nrow(loans) * 100, 2), "%"))
  
table1 <- tibble(status)
knitr::kable(table1, "pipe")
```

In the table, we can tell that around 70% of all loans are paid, and about 28% are in repayment. Therefore, in general, 98% of loans are being paid or already paid. This data could be very encouraging for lenders, as it suggests that most people can get their money back.

#### Loan amount and number of loans across year

Next, we investigate trends over time involving the value and number of loans. From the barchart below, we can see that both the total loan amount and the number of loans issued each year continued to grow until 2012. The highest loan amount borrowed was in the year 2011 at USD 40.2 million. Similarly, the highest of number of loans were taken in the year 2011 as well, at 30,738 loans. It appears that there was a sharp plummet in 2012, however, investigating further we see that the 2012 data only includes values up to April. Further, we verify by locating Kiva's publicly available financial reporting documents, that Kiva's fiscal year ends on December 31th each year. The data itself also contains clues. For all years under `posted_yr`, the corresponding `posted_month` goes from 1 to 12, except for 2012, where `posted_month` stops at 4. This means that the data include loans from January to December from 2005 to 2011, but for 2012, there's only data from January up to April. As a result, the data in 2012 may not reflect the whole year's performance. Hence, we have conclusively determined that the 2012 anomaly is just an artifact of data collection and must be considered in subsequent analysis.

```{r across year,out.width = "80%",header= FALSE, message = FALSE}

# Summarise loan amount by year
year_loan_amount <- loans %>%
  group_by(posted_yr) %>%
  summarise(Total_amount = sum(loan_amount) / 1000) %>%
  mutate(year = factor(posted_yr, posted_yr))

# Summarise loan count by year
year_loan_count <- loans %>%
  group_by(posted_yr) %>%
  summarise(counts = length(unique(loan_id))) %>%
  mutate(year = factor(posted_yr, posted_yr))

# Clean data
year_loan_amount <- na.omit(year_loan_amount)
year_loan_count <- na.omit(year_loan_count)

par(mfrow = c(1, 1))

# Plot loan amount by year
ggplot(year_loan_amount) +
  aes(x = year, y = Total_amount, group = 1, label = Total_amount) +
  geom_line() +
  geom_text(size = 3, nudge_y = 1000) +
  theme(legend.position = "none") +
  xlab("Year") +
  ylab("Total loan amount in USD ('000)") +
  scale_fill_OkabeIto()+
  ggtitle("Distribution of Loan amount over the years")

# Plot loan count by year
ggplot(year_loan_count) +
  aes(x = year, y = counts, group = 1, label = counts) +
  geom_line() +
  geom_text(size = 3, nudge_y = 1000) +
  theme(legend.position = "none") +
  xlab("Year") +
  ylab("Total number of loans") +
  scale_fill_OkabeIto()+
  ggtitle("Distribution of number of loans over the years")
```

####Loan amount and number of loans across country

Next, we consider geographical trends in the Kiva data. Since there are too many countries participating in both lending and borrowing through Kiva, we picked top 10 country with the most transaction based on data in 2011, because 2011 has the most abundant data.

*Geographical Distribution of Borrowers*

```{r loan distribution across borrower country, message = FALSE, warning = FALSE}

# Loan Amount

loan_amount_country <- aggregate(lender_loans$loan_amount, by = list(year = lender_loans$posted_yr, country = lender_loans$location.country), FUN = sum, na.rm = TRUE) %>% 
  rename("loan_amount" = "x") %>% 
  mutate(loan_amount = loan_amount/1000) %>% 
  filter(country %in% c("Peru", "Uganda", "Bolivia", "Cambodia", "Paraguay", "Mexico",
                        "United States", "Tajikistan", "Tanzania", "Azerbaijan"))

# Bar Chart country vs Loan Amount

ggplot(loan_amount_country) +
  geom_line(aes(x=year, y=loan_amount, fill = country, color = country)) +
  xlab("Year") +
  ylab("Total loan count (thousand dollars)") +
  scale_fill_viridis(
  discrete = TRUE,
  option = "D")+
  ggtitle("Top 10 borrower countries by loan amount")

```

From the graph above, we can tell that in half of the top ten borrower countries, Mexico, Paraguay, Tajikistan, Tanzania, and United States, the loan amount has increased from 2005 to 2011, while Uganda, Cambodia, and Peru reached their peak in 2010 and experienced a decrease in loan amount in 2011. Bolivia had the highest loan amount at \$4,500,000 ever in 2009, but the loan amount countinued to drop in 2010 and 2011.

*Geographical Distribution of Lenders*

```{r Lenders Trend, message = FALSE, warning = FALSE, header = FALSE}

country <- aggregate(lender_loans$loan_amount, by = list(year = lender_loans$posted_yr, country = lender_loans$country_code), FUN = sum, na.rm = TRUE)

country_decoded <- country %>%
  rename("country_name" = "country", "loan_amount" = "x") %>%
  filter(country_name %in% c("US", "CA", "GB", "AU", "DE", "NL", "SE", "NO", "JP", "PH")) %>%
  mutate(country_name = countrycode(country_name, "iso2c", "country.name"))

ggplot(data = country_decoded, aes(x = year, y = log(loan_amount), color = country_name)) +
  geom_line() +
  xlab("Year") +
  ylab("Loan amount (log)") +
  guides(color = guide_legend(title = "Country")) +
  scale_fill_viridis(
  discrete = TRUE,
  option = "D")+
  ggtitle("Top 10 lender countries by total loan amount from 2005 to 2012")
```

Again, here in this graph, we picked top 10 countries with the most lender based on 2011. A line graph emphasizes the trend over years. We also took a log for the loan amount. This is because US has a significantly higher total loan amount compared to other countries, due to the fact that Kiva is based in the US. As a result, a log function would make the data more comparable on a line graph. However, when interpretting the line graphs, keep in mind that log function also reduce the absolute magnitude of increase/decrease of the changes in lines. All lines peak in 2011, and drop intensely in 2012, which matches the result presented in the last two graphs.

### -- Part 1: Lender Motivation --

Kiva provides a platform where borrowers post their credit needs, and lenders choose how much and to whom they lend; this part of the prompt asks us to investigate what motivates lenders.

The most direct way to investigate lender preferences is to utilize the variable `loan_because`.

While the `loan_because` variable allows us to look directly at lenders self-reported motivation, analysis of this variable requires more advanced text processing which we will discuss in a later section. As a next-best option, we know look at the characteristics of borrowers in hopes of identifying patterns in lender preferences.

```{r Loan_because, warning = FALSE}
head(unique(lenders$loan_because))
```

An alternate way of approaching this part of the prompt is to attempt to infer lender motivation from lender and borrower characteristics. In this vein, one interesting question is which borrower characteristics impact their success at obtaining funding; using this line of thinking, we might try to infer lender motivation from the differences between funded and non-funded projects. Similarly, we might investigate how manifest lender preferences vary with lender characteristics.

```{r Lenders top borrowers table}
top_lenders <- lender_loans %>%
  filter(country_code %in% c("US", "CA", "GB", "AU", "DE", "NL", "SE", "NO", "JP", "PH"))

top_lender_borrow <- aggregate(top_lenders$loan_amount, by = list(lender_country = top_lenders$country_code, borrower_country <- top_lenders$location.country), FUN = sum, na.rm = TRUE) %>%
  mutate(lender_country = countrycode(lender_country, "iso2c", "country.name")) %>%
  arrange(desc(x)) %>%
  arrange(lender_country) %>%
  group_by(lender_country) %>%
  slice(1, 2, 3) %>%
  mutate(x = format(x, big.mark = ",", trim = TRUE)) %>%
  rename("Loan Amount (USD)" = "x", "Lender Country" = "lender_country", "Borrower Country" = "Group.2")

# table_org <- t(top_lender_borrow)
# 
# for (i in 4:30)
#   +{if (i%%3 == 1)
#   {
#     rows <- nrow(table_org)
#     table_org[rows+1, ]<- NA
#     rownames(table_org)[rownames(table_org) == "i"] <- make.names("Lender Country", unique = FALSE)
#   }}

knitr::kable(top_lender_borrow, "pipe")
```

In the table above, we can see the top three recipient countries of each of the top ten lender countries. This graph shows that lenders may not necessarily prefer borrowers from their own countries.

THIS TABLE IS GOOD BUT IT'S TOO LONG. INSTEAD OF LISTING THE TOP THREE BORROWERS FOR EACH COUNTRY VERTICALLY, DO IT HORIZONTALLY SO THAT THERE IS ONE ROW FOR AUSTRALIA AND A COLUMN FOR TOP BORROWER, SECOND, THIRD. AND THEN DROP THE LOAN AMOUNT

### -- Part 2: Repayment Factors --

Now let's switch to part 2 of the prompt, repayment factors.

#### Sector

The borrower's sector could have a strong impact on both the demand for loans and the ability of borrowers to repay their loans. Some sectors could need more liquid capital than others, thus making up a bigger proportion of loan requests on Kiva.com. On the other hand, lenders might have a preference for lending to certain sectors, driven either by personal priorities or a belief that some sectors may be more profitable than others, and hence more likely to repay their loans.

```{r Sector, Paid amount, message=FALSE, warning=FALSE}
sector_loan <-
  aggregate(lender_loans$loan_amount, by = list(sector = lender_loans$sector), FUN = sum)
sector_paid <-
  mutate(aggregate(lender_loans$paid_amount,
    by = list(sector = lender_loans$sector),
    FUN = sum, na.rm = TRUE
  ))

by_sector <- sector_loan %>%
  mutate(x = x/1000) %>% 
  mutate(paid = sector_paid$x/1000) %>%
  mutate(unpaid = (x - paid)) %>%
  rename(loan = x)

by_sector <- by_sector[order(by_sector$loan, decreasing = TRUE), ]
by_sector_paid <- select(by_sector, sector, paid) %>%
  mutate(status = "paid") %>%
  rename("amount" = "paid")
by_sector_unpaid <- select(by_sector, sector, unpaid) %>%
  mutate(status = "unpaid") %>%
  rename("amount" = "unpaid")
by_sector <- rbind(by_sector_paid, by_sector_unpaid) %>%
  group_by(sector)

ggplot(by_sector) +
  aes(x = reorder(sector, amount), y = amount, fill = status, label = amount) +
  geom_bar(stat = "identity", position = "stack") +
  theme() +
  coord_flip() +
  xlab("Sector") +
  ylab("Loan amount (thousand dollars)") +
  scale_fill_viridis(
  discrete = TRUE,
  option = "D")+
  ggtitle("Paid vs Unpaid Loan Amount by Sector")
```

The graph above shows that for most sectors, paid loans well exceed unpaid loans. Of the top five sectors in terms of loan amount, Agriculture has the highest proportion of unpaid loans, around 30%. Of all sectors, Housing and Education have the highest unpaid loans, nearly 50%. Arts has the lowest proportion of unpaid loans, about 18%. Whether or not sector is an important factor that predicts the the borrowers' ability to repay the loans is thus meaningful to investigate more deeply.

```{r Sector loan amount with gender, message = FALSE, warning = FALSE }

# Get unique borrower id and total male female borrower count
loan_sub <- loans %>%
  group_by(borrower_image_id) %>%
  summarize(total_m = sum(borrower_m_count), total_f = sum(borrower_f_count))

# Create new borrower type column based on conditions
loan_sub <- loan_sub %>%
  mutate(bor_type = if_else(total_m == 1 & total_f == 0, "Male",
    if_else(total_m == 0 & total_f == 1, "Female", "Group")
  ))

pc_gender <- subset(loan_sub, loan_sub$bor_type %in% c("Male","Female"))

merge <- left_join(x = lender_loans, y = pc_gender, by = "borrower_image_id")

sec_gender_loan <- merge %>%
  group_by(sector, bor_type) %>%
  summarise(tot_amount = sum(loan_amount) / 1000) %>%
  arrange(desc(tot_amount))

text <- aggregate(sec_gender_loan$tot_amount, by = list(gender = sec_gender_loan$bor_type), FUN=sum)

cat("Female proportion:", text$x[1]/(text$x[1]+text$x[2]), "Male proportion:", text$x[2]/(text$x[1]+text$x[2]))
  
par(mfrow = c(1, 1))

sec_gender_loan <- na.omit(sec_gender_loan)

ggplot(sec_gender_loan) +
  aes(x = reorder(sector, tot_amount), y = tot_amount, fill = bor_type, label = tot_amount) +
  geom_bar(stat = "identity", position = "stack") +
  theme() +
  guides(fill=guide_legend(title="Borrower Composition"))+
  coord_flip() +
  xlab("Sector") +
  ylab("Total loan amount (thousand dollars)") +
  scale_fill_viridis(
  discrete = TRUE,
  option = "D")+
  ggtitle("Sector vs Loan Amount by Gender")
```



Next, we investigate gender within each sector. Overall, the total loan amount of female borrowers (58.54%)  exceeds that of male borrowers (41.45%). Similarly, in most sectors there is a greater proportion of female borrowers. Nevertheless, in Transportation, Construction, and Manufacturing, there are males received a higher loan amount than females. In Agriculture and Services, the proportion of loan amounts received by males and females appear to be close.

Compared to the Paid vs Unpaid Loan Amount graph, we can see that although Agriculture has the highest unpaid loan proportion, the gender composition is about balanced. Although females dominate Housing, Education, and Arts, the paid and unpaid proportions appear to be very different. Therefore, we may conclude that gender is not related to whether or not the loan can be repaid.

### Conclusion

In this section, we observed several interesting trends in regards to lender motivation and loan repayment. In particular, we noticed variability in `paid` vs unpaid proportion in terms of `sector`. This could reflect a potential predictor of the borrower's ability to repay, yet it can also be due to the sheer difference in total loan amount in different sectors, as well as the economic structure in each of the borrowers' countries. Again, EDA is the very first step to observe superficial correlation in the data, and this tutorial only shows some of the options. You can also experiment other variables or other visualization presentations, which may inspire you to a completely different direction that could catch the judges' eyes!

For further more in-depth analysis, it's definitely worthy to test whether the discrepency discovered in variables in this section is statistically significant. For this purpose, we can apply statistical models, such as regression models. We can also do text analysis to look at different types of data apart from the numeric data we explored in this section.
