---
title: "Analysing Text Data to understand Lender Motivation"
author: "Aarushi Verma"
date: "01/07/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

------------------------------------------------------------------------

## Introduction

In this section we continue our analysis of part one of the prompt: motivations for lending. In the previous section we attempted to infer lender motivation from systematic differences in funded and non-funded borrowers, here we directly analyze the free responses provided by lenders using text analysis. Text analysis is a machine learning technique where text data is parsed and structured to extract insights that are valuable and meaningful. It's purpose is to gain information from free text content. 

Our data contains two relevant columns containing free text quoting the self-reported lenders' reasons for lending and borrowers' reasons for their intended use of the loan amount. We will perform analysis on those text fields and assess what may have made a lender favor a certain borrower. By looking at lender's reasons we can directly infer what motivates them. We also compare the borrowers' intended uses of the loan amount for "funded" and "not funded" borrowers to identify if there are specific words used by funded borrowers that  motivate a lender to lend them money.

## Setup

### Importing required libraries

Before we dive into the analysis, we need to import all the required packages. In this analysis, the packages we have used are:

```{r libraries, warning=FALSE, echo=TRUE, message=FALSE}
#text analysis packages
library("tm")
library("tidyverse")
library("textstem")

#visualization packages
library("wordcloud")
library("ggplot2")
library("plotly")
library("colorblindr")
library("viridis")

library("rmarkdown")
library("readr")
library("kableExtra")
options(readr.show_col_types = FALSE)
```

### Reading the data

We use the `lender_loans.csv` file for our analysis. The columns of interest to us are:

-   `loan_because` : free response indicating lender's reason for lending
-   `use` : free response indicating a borrower's intended use (more flexible than "sector")

*The data set presented below is a subset of columns from the actual data*

```{r input_data, echo=TRUE, warning=FALSE, results='asis', message = FALSE}

kiva_ll <- read_csv("data/lender_loans.csv")
data_preview <- subset(kiva_ll, select=c("loan_id","loan_because","use","loan_amount"))

#Specifying rows for which we have data to showcase various reasons provided by lenders for lending
X <- (data_preview[c(133,165,177,235,298),]) 

knitr::kable(X, "pipe")
```

Before we proceed with the analysis, we must check the proportion of missing values in our data. 

```{r missing, echo=TRUE, warning=FALSE, results='asis', message = FALSE}

lender_missing <- sum(is.na(kiva_ll$loan_because))/nrow(kiva_ll)
bor_missing <- sum(is.na(kiva_ll$use))/nrow(kiva_ll)

sprintf("The proportion of missing values in loan_because column are %f ",round(lender_missing,2))
sprintf("The proportion of missing values in use column is %f ",round(bor_missing,2))

```

TO UPDATE ASSUMPTIONS FOR DATA MISSINGNESS AND WHETHER IT IS MISSING AT RANDOM OR NOT


## Analysis

In this section, we process the aforementioned free text fields to understand lender motivation. We convert the text fields into corpora, perform data cleaning and some advanced text transformations. Further, we visualize this data to extract our insights. We analyse the text from both lenders and borrowers perspectives. From the lenders' text field we identify the most common words used by lenders when citing reasons for lending. From the borrowers' text field, the most common words used can give us an insight into what the loans are typically used for. We further try to separate these loans into funded and not funded to assess whether there is a trend in the words used by borrowers' whose loans got funded versus words used by borrowers whose loans did not get funded. 


### Lender text analysis

------------------------------------------------------------------------

#### Creating a corpus

Now that we've identified the relevant data, the first step towards text analysis is to clean the data and convert it to a corpus - a format used for text analysis, and the format most convenient for the $\texttt{tm}$ package which we will be using.

Upon first inspection, we observe multiple special characters present which are not recognized by the typical `UTF-8` encoding of the file. Therefore we first encode the field using `latin1` encoding so that it can recognize these special characters.

We also observe that each line in our data frame does not represent a unique lender or their reason for lending. If a specific lender lends to multiple borrowers, their reason for lending will reappear as many times in our data as well. This will sway the frequency of the words used. For the purpose of this analysis we will only count each lender once therefore, capturing their reasons for lending once as well. 

We subset our data frame to only contain the unique `lender_ids` and their corresponding reasons for lending as captured in the column `loan_because`.

```{r corpus, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# To recognize special characters
Encoding(kiva_ll$loan_because) <- "latin1" 


loan_bec <- subset(kiva_ll, select=c("lender_id","loan_because"))
loan_bec <- unique(loan_bec[ , c("lender_id", "loan_because")])

# Convert input data to corpus
corpus <- Corpus(VectorSource(as.vector(loan_bec$loan_because))) 
```
<br/>

#### Data Transformation

Once we have created our corpus, we transform our data via the following steps:

1.  Removing all special characters and punctuation

2.  Converting all text to lower case

3.  Removing all numbers

4.  Removing extra white spaces

5.  Removing "stop words." Stop words refer to common words used in text which do not add any contextual value. For example: 'a','the','as', etc.

By performing these steps, we are able to get rid of the 'noise' in the data. For example, by converting to lowercase it helps to keep the data consistent for analysis. By removing stop words it helps to focus mainly on key words present in the text which are relevant to us.

```{r cleaning, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
#Define a function to replace characters with space
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))

corpus <- tm_map(corpus, toSpace, "/")
corpus <- tm_map(corpus, toSpace, "@")
corpus <- tm_map(corpus, toSpace, "\\|")

#Convert all text to lower case
corpus <- tm_map(corpus, content_transformer(tolower))

# Remove  numbers
corpus <- tm_map(corpus, removeNumbers)

# Remove english common stopwords
corpus <- tm_map(corpus, removeWords, stopwords("english"))

# Remove punctuation
corpus <- tm_map(corpus, removePunctuation)

# Eliminate extra white spaces
corpus <- tm_map(corpus, stripWhitespace)
```
<br/>

#### Lemmatization

Lemmatization is a method used in text analysis to remove the inflections and retain only the root or dictionary form of the word. This dictionary form of the word is referred to as the *lemma*. This is done by using morphological analysis. The function `lemmatize_strings` uses a lookup dictionary (Mechura’s (2016) English lemmatization list available from the lexicon package) and reduces the strings to their lemma forms. Another method used in text analysis is Stemming. Stemming works by cutting off the affixes. The algorithm looks at common prefixes and suffixes that may occur for a word and reduce it to its stem. However, stemming presents some limitations since the results it presents do not retain the context of a word as lemmatization does and the results may  not desirable. Look at the following vector as an example. The vector x represents different examples of the word purchase. We see that when we stem, simply the sufix of the words is reduced to the root word *purchas*. Whereas, when we use lemmatization, our result is the lemma of the word and context is retained when the word is *purchaser*. 
```{r}
x <- c('purchase', 'purchased', 'purchaser', 'purchasing', 'purchases')

stem_words(x)
lemmatize_words(x)

```

```{r lemming, echo=TRUE, results='hide', message=FALSE, warning=FALSE }
#Lemmatization
corpus <-tm_map(corpus, lemmatize_strings)
```

<br/>

#### Word Frequency Matrix

Now that the data is prepared, we begin our analysis with a word frequency matrix, which displays the frequency of occurrence for each word in our data.

For this analysis we consider only the raw count of each term, however alternative weighting and normalization schemes are possible in the package as well. We convert the matrix to a data frame for ease of manipulation and sort it in terms of decreasing frequency.

```{r matrix, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Build a term-document matrix
TextDoc_dtm <- TermDocumentMatrix(corpus)

# Convert data type to matrix
dtm_m <- as.matrix(TextDoc_dtm)

# Sort by descending value of frequency
dtm_v <- sort(rowSums(dtm_m),decreasing=TRUE)

#Convert to dataframe
dtm_d <- data.frame(word = names(dtm_v),freq=dtm_v)
head(dtm_d)

```
<br/>

#### Visualizing the data

**Bar Chart of top 10 most frequent words**

Once we have our word frequency data frame we can visualize it. First we plot a bar chart to see the 10 most frequent words used by lenders along with their frequency. From our bar chart we see that the most commonly used word is *help*, with a frequency of 869, which is nearly twice that of the next most commonly used word, *people* . This suggests that an altruistic desire to help entrepreneurs may be a significant motivation for Kiva lenders.

```{r barchart, echo=TRUE, fig.align='center',warning=FALSE}
# Plot the most frequent words

ggplot(dtm_d[1:10,])+ aes(x = reorder(word, -freq), y=freq)+
       geom_bar(stat="identity", fill = "blue", alpha=.6, with = 0.5)+
  xlab("Word")+
  ylab("Frequency")+
  theme(axis.text.x=element_text(size=12))+
  ggtitle("Top 10 most frequent words [Lenders]")

```

**Word Cloud**

A word cloud is visual representation of words in the form of a cluster where the size of the word represents its frequency. For our word cloud, we plot all the words in the data set which have a minimum frequency of 5. While the interpretation is similar to the that of the bar chart, this alternative visualization allows us to see more words than could reasonably fit in a bar chart.

```{r wordcloud, echo= TRUE, fig.align='center'}
set.seed(1234)
wordcloud(words = dtm_d$word, freq = dtm_d$freq, min.freq = 5,
          max.words=100, random.order=FALSE, rot.per=0.40,scale=c(4,0.75),
          colors=brewer.pal(8, "Dark2"))
```

This concludes our analysis of the `loan_because` column. We looked at our data from the Lender's perspective and observed that most common word used is *help*, as is evident from the bar chart as well. However, from the word cloud we can also observe that most of the words used by lenders seem to be in the sphere of helping people, indicating that lenders sign up on Kiva provide support and be of service to entrepreneurs. The difference in the frequency of the words also indicate a heterogeneity in lender's motivation. Words such as *bless*, *can*,*able* and *fortunate* suggest that lender's are motivated by the abundance of resources at their disposal and are driven to share these resources. Other words such as *empower*, *community*, *succeed* and *achieve* indicate a motivation that stems from a larger picture of society. The choice of words implies their motivation to enable communities to increase control over their lives by helping them become more independent.


### Borrower text analysis

------------------------------------------------------------------------

Next, we look for systematic differences in the words used by funded and non-funded borrowers when describing their intended use for the loans. This may suggest lender preferences and motivations. In this portion of the analysis, we rely on the `status` column, which indicates the current status of the loan, as well as the `use` column, which indicates the borrower's intended use for the requested loan amount.

To get a better idea of how many loans are funded, we look at the unique values in the `status` column and the number of loans associated with each value. Further, we use these values to create an indicator variable for whether a loan has been funded or not.

```{r status, echo= TRUE, messgae = FALSE}

#Looking at unique values and respective loan count in status column
status_table <- kiva_ll %>% 
  group_by(status) %>% 
  summarize(count = length(unique(loan_id))) %>% 
  arrange(desc(count))

status_table %>% kbl(caption = "Value counts of status column") %>% 
  kable_classic(bootstrap_options ="striped",full_width = F, html_font = "helvetica",position = "left")
```

The status column contains 7 unique values. We look at the data dictionary and Kiva website to get a better idea of what each term means. Based on that we make the classification of these terms to our 2 classes funded and not-funded:

-   Funded: $\texttt{funded, in_repayment, paid, defaulted, refunded}$

-   Not Funded: $\texttt{expired, fundraising}$


For funded we included status values which indicate that a loan has been funded already, irrespective of what the current status of the loan is. For not funded we included the status values which indicate that a loan was never funded or is in the process of being funded.

```{r funding indicator,echo= TRUE, message = FALSE}

#creating a vector for values which are associated with funded loans
funded_vec <- c("paid","in_repayment","funded","defaulted", "refunded")

#create indicator variable for Funded or not funded loans
funding_stat <- kiva_ll %>% 
  mutate(funding_status = if_else(kiva_ll$status %in% funded_vec, "Funded","Not Funded"))

#Create dataframe with unique borrower id, use for loan and corresponding funding status
loan_use <- unique(funding_stat[ , c("borrower_image_id", "use","funding_status")])
```

We filter our dataset to these values and create our `funding_status` indicator variable. Next, we separate the data into two data frames - Funded loans and Not funded loans. We then perform the same text analysis steps as we did in the lenders section on the `use` column.

```{r borrower cleaning, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
Encoding(loan_use$use) <- "latin1" #To recognize special characters

funded_loan <- subset(loan_use,loan_use$funding_status == "Funded")
notfunded_loan <- subset(loan_use,loan_use$funding_status == "Not Funded")

funded_corpus_use <- Corpus(VectorSource(as.vector(funded_loan$use)))# Convert input data to a corpus
notfunded_corpus_use <- Corpus(VectorSource(as.vector(notfunded_loan$use)))

funded_corpus_use <- tm_map(funded_corpus_use, toSpace, "/")
funded_corpus_use <- tm_map(funded_corpus_use, toSpace, "@")
funded_corpus_use <- tm_map(funded_corpus_use, toSpace, "\\|")

notfunded_corpus_use <- tm_map(notfunded_corpus_use, toSpace, "/")
notfunded_corpus_use <- tm_map(notfunded_corpus_use, toSpace, "@")
notfunded_corpus_use <- tm_map(notfunded_corpus_use, toSpace, "\\|")


#Convert all text to lower case
funded_corpus_use <- tm_map(funded_corpus_use, content_transformer(tolower))
notfunded_corpus_use <- tm_map(notfunded_corpus_use, content_transformer(tolower))

# Remove  numbers
funded_corpus_use <- tm_map(funded_corpus_use, removeNumbers)
notfunded_corpus_use <- tm_map(notfunded_corpus_use, removeNumbers)

# Remove english common stopwords
funded_corpus_use <- tm_map(funded_corpus_use, removeWords, stopwords("english"))
notfunded_corpus_use <- tm_map(notfunded_corpus_use, removeWords, stopwords("english"))

# Remove punctuation
funded_corpus_use <- tm_map(funded_corpus_use, removePunctuation)
notfunded_corpus_use <- tm_map(notfunded_corpus_use, removePunctuation)

# Eliminate extra white spaces
funded_corpus_use <- tm_map(funded_corpus_use, stripWhitespace)
notfunded_corpus_use <- tm_map(notfunded_corpus_use, stripWhitespace)

#Lemmatize strings
funded_corpus_use <-tm_map(funded_corpus_use, lemmatize_strings)
notfunded_corpus_use <-tm_map(notfunded_corpus_use, lemmatize_strings)
```

```{r use matrix, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# Build a term-document matrix
funded_TextDoc_dtm <- TermDocumentMatrix(funded_corpus_use)
notfunded_TextDoc_dtm <- TermDocumentMatrix(notfunded_corpus_use)

# Convert data type to matrix
funded_dtm_m <- as.matrix(funded_TextDoc_dtm)
notfunded_dtm_m <- as.matrix(notfunded_TextDoc_dtm)

# Sort by descending value of frequency
funded_dtm_v <- sort(rowSums(funded_dtm_m),decreasing=TRUE)
notfunded_dtm_v <- sort(rowSums(notfunded_dtm_m),decreasing=TRUE)

#Convert to dataframe
funded_dtm_d <- data.frame(word = names(funded_dtm_v),freq=funded_dtm_v)
notfunded_dtm_d <- data.frame(word = names(notfunded_dtm_v),freq=notfunded_dtm_v)

```

**Bar Chart**

First we plot two bar charts to visualize and compare the top 10 most frequently used words by borrowers whose loans have been funded and not funded.

```{r borrower barchart, echo=FALSE, warning=FALSE, fig.show='hold'}
# Plot the most frequent words

par(mfrow=c(1,1)) 

#Borrower Bar Chart of top 10 frequent words
ggplot(funded_dtm_d[1:10,])+ aes(x = reorder(word, -freq), y=freq)+
       geom_bar(stat="identity", fill = "blue", alpha=.6,with = 0.5)+
  xlab("Word")+
  ylab("Frequency")+
  coord_flip()+
  theme(axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))+
  ggtitle("Top 10 most frequent words in funded loans")

ggplot(notfunded_dtm_d[1:10,])+ aes(x = reorder(word, -freq), y=freq)+
       geom_bar(stat="identity", fill = "blue", alpha=.6,with = 0.5)+
  xlab("Word")+
  ylab("Frequency")+
  coord_flip()+
  theme(axis.text.x=element_text(size=12), axis.text.y=element_text(size=12))+
  ggtitle("Top 10 most frequent words in not funded loans")

```


On comparing these bar charts it is interesting to see that the top 3 words for both funded and not funded loans are the same - *buy*, *purchase* and *business*. he top 2 words also have the approximately more than double the frequency of the others. This implies that most borrowers are applying for loans to be able to source the raw materials for their businesses. We must also take into consideration the difference in the number of funded and not funded loans in our data. The number of loans affects the borrowers' quotes in the `use` column which therefore affects the number of times a word will appear in the data.

Among the remaining words we see words like *rice* and *fertilizer* and *supply* in funded loans whereas these are not present in the not funded loans. These words indicate specificity in borrowers' intended use of the loan amount which could imply that lenders look for borrowers with a concrete plan. For not funded loans we see the words *food*, *additional* and *pay* which are not present in the funded loans. This could reiterate that lender's are motivated by specificity and not by general terms used by borrowers when quoting reasons for requesting the loan amount. 

**Word Cloud**

Similar to the bar charts we create two word clouds, one representing the most common words used by the borrowers whose Loans have been funded and the other representing loans that did not get funded. We plot all the words in the data set which have a minimum frequency of 5. While the bar chart displays the top 10 most frequent words, the word cloud gives us a more holistic view of the words used by borrowers when requesting for loans. 

```{r, echo=TRUE, header=FALSE, message = FALSE, }
#Table to see no. of funded and not funded loans

funding_loans_count <- funding_stat %>% group_by(funding_status) %>% summarize(loan_count = length(unique(loan_id)))

funding_loans_count %>% kbl(caption = "Value Counts of funding status",col.names=c("Funding status", "No. of loans")) %>% 
  kable_classic(bootstrap_options ="striped",full_width = F, html_font = "helvetica",position = "left")

```

```{r, echo=TRUE,header= FALSE, message = FALSE, warning = FALSE}
#Borrower Wordcloud - Funded Loans
set.seed(1234)


layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Wordcloud for Funded Loans")
wordcloud(words = funded_dtm_d$word, freq = funded_dtm_d$freq, min.freq = 5, main ="Wordcloud [Borrowers]",random.order=FALSE, rot.per=0.40,scale=c(4,0.75),
          colors=brewer.pal(2, "Dark2"))
```

```{r, echo=TRUE,header= FALSE, message = FALSE, warning = FALSE}
#Borrower Wordcloud - Not Funded Loans
set.seed(1234)

layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Wordcloud for Not Funded Loans")
wordcloud(words = notfunded_dtm_d$word, freq = notfunded_dtm_d$freq, min.freq = 5, main ="Wordcloud [Borrowers]",random.order=FALSE, rot.per=0.40,scale=c(4,0.75),
          colors=brewer.pal(6, "Dark2"))

```

The biggest difference in the wordclouds between funded and not funded loans is the variety of words depicted in the funded loans word cloud. As mentioned earlier (and depicted in the table) the difference in funded and not funded loans is huge. This offers a larger variety of words among funded loans. The size of the words in the wordcloud represent its frequency in the data. We see that the variation in sizes in the funded wordcloud is also limited indicating that there is more variety in the words used rather than certain words used which may motivate lenders. For not funded loans, we see fewer words but they do share similarity to words present in the funded loans word cloud as well. However, we do see a lot more variation in terms of the frequency of these words. This suggests that a lender's motivation is not solely affected by the borrower's reason for requesting a loan. There are other factors a lender is taking into consideration which affect their motivation to a larger extent than the words used by a borrower.


## Conclusion

In this tutorial, we analyzed free text fields containing quotes from both lenders and borrowers to understand what motivates lenders. We observed that based on the text analysis of lenders' self reported reasons for lending they tend to be motivated by altruism and most commonly lend money to help entrepreneurs. When it comes to borrowers, we observe that most borrowers seek out loans in order to source raw materials or to purchase resources for their business or workplace. We also looked at a breakdown of Funded and Not funded loans for borrowers. While the differences are not stark between funded and non-funded borrowers, we observe that there is a lot more variety in the words used by funded borrowers. It suggests that lenders are more interested in unique business ideas and plans. However, the words in both word clouds also share common words implying that lender's motivation is affected by other factors such as the loan amount, the sector the loan is for and whether the borrower has been delinquent in the past or not, to name a few.
