---
title: "Hypothesis Testing "
author: "Aarushi Verma"
date: '2022-07-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this section we focus on the first part of the prompt, which asks us to analyze the motivations for lending among contributors to Kiva. Our approach here is suitable for students with limited statistical background: we first explore the general characteristics of borrowers, then we contrast the characteristics of funded borrowers to those of non-funded borrowers, using both descriptive statistics and hypothesis testing. Since lenders choose whom they fund, the results of this analysis will reveal the types of borrowers which lenders prefer, and may suggest motivations for lending.

An important caveat to this analysis is that the 2012 DataFest dataset is not an exhaustive record of all Kiva borrowers, but a subset selected for the purposes of the competition. In order to proceed with our analysis, we must assume that the data provided is a representative sample. However, note that if this assumption is violated, i.e., if the data differs systematically from the full population of Kiva transactions due to seleciton bias, our inferences will not be valid.

THIS CAVEAT SHOULD BE IN THE EDA OR GETTING STARTED SECTION TOO AND JUST REFERENCE IT AGAIN HERE AND EVERYWHERE ELSE.

## Setup

Before we dive into the analysis, we need to import all the required packages. This includes packages for statistical analysis and visualization - many of which you will be familiar with from earlier sections.

### Importing the required packages

The packages we have used in this walk through are:

```{r libraries, warning=FALSE, echo=TRUE, message=FALSE}
library("ggplot2")
library("plotly")
library("tidyverse")
library("readr")
library("colorblindr")
library("viridis")
library("memisc")
library("kableExtra")
options(readr.show_col_types = FALSE)
```

### Reading the data

We use the `lender loans` and `loans` files for this analysis. We import these files using the `read_csv` function.

WHY, WHAT'S IN THESE FILES

```{r input_data, echo=TRUE, warning=FALSE, results='asis'}

kiva_ll <- read_csv("main files/lender_loans.csv")
kiva_loan <- read_csv(("main files/loans.csv"))
```

## Analysis

### Borrower Characteristics

We begin by getting an overview of the characteristics of all would-be borrowers, including their gender, sector, country, and requested loan amount. Because we will ultimately be interested in differences between funded and non-funded borrowers, we first add an indicator for borrower type and funding status to our data set. This will help improve our visualizations and communicate trends at these two levels.

**Borrower Type**

DEFINE BORROWER TYPE

We are currently using the `lender_loans.csv` file which only contains the total number of borrowers. However, the `loans.csv` file contains a split of the count of female and male borrowers. Using this information we can create a new variable which indicates whether a borrower is an individual male borrower, individual female borrower or part of a group. We use the `mutate` function to create this new columnm and name it `bor_type`. We update the values based on `if_else` conditions. We apply these conditions on the male borrower and female borrower count column. The conditions we use are:

-   If the value in the m_borrower_count column is equal to 1 and f_borrower_count is equal to 0, we update the value in the `bor_type` column as *Male*

-   If the value in the f_borrower_count column is equal to 1 and m_borrower_count is equal to 0, we update the value in the `bor_type` column as *Female*

-   For any other case, we update the value in the `bor_type` column as *Group*

```{r borrower type variable , echo=TRUE,warning=FALSE,results='hide', warnings = FALSE, message=FALSE}
#Get unique borrower id and total male female borrower count
kiva_loan_sub <- kiva_loan %>% group_by(borrower_image_id) %>% 
  summarize(total_m = sum(borrower_m_count),total_f = sum(borrower_f_count))

#Create new borrower type column based on conditions
kiva_loan_sub <- kiva_loan_sub %>%
  mutate(bor_type = if_else(total_m == 1 & total_f == 0, "Male",
                       if_else(total_m == 0 & total_f == 1, "Female","Group")))
```

After creating the indicator variable, we merge those values to the `lender_loans` data frame and check for NAs. This will help us verify if all the values in our left data frame have found corresponding values or not. For the purpose of this analysis we will drop any NAs in the data since it is not possible to obtain that information elsewhere.

```{r master df creation, echo = TRUE, warning=FALSE, results='hide', message=FALSE}

#Checking if each borrower image id maps to 1 Gender or group
kiva_loan_sub %>% 
  group_by(borrower_image_id) %>% 
  summarize(count=n_distinct(bor_type)) %>% 
  arrange(desc(count))

#Merging sex group to main data
kiva_ll_merged <- left_join(x = kiva_ll, y = kiva_loan_sub, by="borrower_image_id")

# Create df for plotting
bor_chr <- kiva_ll_merged %>% 
  group_by(borrower_image_id, loan_id, location.country, sector,bor_type) %>%
  summarize(tot_loans = sum(loan_amount), borr_count = max(borrower_count)) %>% 
  arrange(desc(tot_loans))
```

**Funding Status**

Our data contains a `status` column which indicates the current status of the loan. First we look at the unique values in this column and the number of loans associated with each respectively. Further, we use these values to create an indicator variable for whether a loan has been funded or not.

```{r status unique, echo= TRUE, messgae = FALSE, warning=FALSE}

#Looking at unique values and respective loan count in status column
status <- kiva_ll_merged %>% group_by(status) %>% summarize(count = length(unique(loan_id)))

status %>% kbl(caption = "Value Counts of Status column") %>% 
  kable_classic(bootstrap_options ="striped",full_width = F, html_font = "helvetica",position = "left")
```

<br/> The status column contains 7 unique values. We look at the data dictionary and Kiva website to get a better idea of what each term means. Based on that we make the classification of these terms to our 2 classes funded and not-funded:

-   Funded: $\texttt{funded, in_repayment, paid}$

-   Not Funded: $\texttt{expired, fundraising}$

-   We do not include the loans associated with the status $\texttt{defaulted}$ and $\texttt{refunded}$ in our analysis since they do not clearly fall in either of the classes we have created

```{r funding indicator,echo= TRUE, message = FALSE, warning = FALSE, include=TRUE}

#creating a vector to subset data for values corresponding to funded and not funded
funding <- c("expired","funded","fundraising","in_repayment","paid")

#subset data to values in status column that belong to funding vector.
kiva_ll_funding<- subset(kiva_ll_merged,kiva_ll_merged$status %in% funding)

#creating a vector for values which are associated with funded loans
funded_vec <- c("paid","in_repayment","funded")

#create indicator variable for Funded or not funded loans
kiva_ll_funding <- kiva_ll_funding %>% mutate(funding_status = if_else(
                       kiva_ll_funding$status %in% funded_vec, "Funded","Not Funded"))

```

Once we have added our new indicator columns, we will use descriptive statistics to visualize these and identify trends, if any to gain some insight into borrower characteristics

We look at the distribution of the funding status of loans among borrower types for various sectors. We will visualize this data in a tabular format since our data contains two categorical variables.

```{r contingencytable, echo = TRUE, warning = FALSE, message = FALSE, results='asis'}

#Creating master df for all plots
bor_chr <- kiva_ll_funding %>% group_by(borrower_image_id,loan_id,location.country,sector,bor_type, funding_status) %>%
  summarize(tot_loans = sum(loan_amount),borr_count = max(borrower_count)) %>% arrange(desc(tot_loans))


# Creating table to see number of loans and average loan amount given to a borrower for each sector further split between funded and not funded

mytable <- table(bor_chr$sector, bor_chr$funding_status, bor_chr$bor_type)
Y <- ftable(mytable, col.vars = c(2,3))
knitr::asis_output(htmltools::htmlPreserve(
   format_html(Y,
               show.titles=TRUE,
               digits=0,
               format="f",
               toprule=2,midrule=1,bottomrule=2,
               split.dec=TRUE,
               style = ftable_format_stdstyle,
               margin="2ex auto", 
               varontop,varinfront)
))
```

TOO MUCH INFO IN THIS TABLE, I WOULD NOT SEPARATE BY FUNDING STATUS (THAT WILL BE YOUR LATER ANALYSIS) AND PROBABLY CHANGE TO SIDE BY SIDE BAR GRAPHS WITH X-AXIS SECTOR AND TWO BARS FOR EACH SECTOR (MALE AND FEMALE)

This table gives us a snapshot of how the number of loans are distributed among sectors. The food sector has the highest number of loans that have been funded specifically for individual female borrowers. Interestingly, the highest number of loans that have not been funded also belong to the Food sector. These fall into the individual male borrower category.

To further dive into our borrower characteristics, we visualize the data from various aspects in our data.

First, we look at how the average loan amount per borrower or borrower group varies across countries within each sector. Our data contains 64 unique countries that a borrower may belong to. Visualizing all the countries on a graph would make it too crowded and illegible. For the purpose of this analysis, we filter our data to the top 10 countries with the highest total loan amount. The bar chart below represents the top 10 countries. ( *This is irrespective of funding status* )

```{r top 10 country , echo = TRUE, include= TRUE, results ='asis'}
# Loan Amount Calculation
tot_loans_country <- bor_chr %>% group_by(location.country) %>%
        summarize(tot_loans = sum(tot_loans)/1000) %>%
  arrange(desc(tot_loans)) %>%
  mutate(country = factor(location.country, location.country))

#Filter top 10 countries with highest loan amount
top_10_tot_loan <- head(tot_loans_country,10)

#Bar Chart country vs Loan Amount
ggplot(top_10_tot_loan)+
       geom_bar(aes(country,tot_loans, fill =country), stat="identity")+
  theme(legend.position = "none",axis.text.x = element_text(angle = 30))+
    xlab("Country")+
  ylab("Total loan amount in USD ('000s)")+
  scale_fill_viridis(
  discrete = TRUE,
  option = "D") +
  ggtitle("Top 10 countries by total loan amount")

```

IT'S NOT CLEAR TO ME HOW THE BY-COUNTRY ANALYSIS RELATES TO THE PROMPT. I WOULD REDUCE OR REMOVE THIS.

THERE IS TOO MUCH INFORMATION IN ALL OF THESE GIANT FACETED BAR PLOT, LET'S TALK ABOUT WHAT YOU WANT TO COMMUNICATE AND FIGURE OUT HOW TO DO IT MORE CLEARLY.

The rest of our analysis is limited to these 10 countries. We filter our data to these countries, calculate the average loan amount and perform our descriptive statistics.

First, we look at how the average loan amount varies across countries, for each borrower type. Since the number of borrowers is not evenly distributed across types in our data, looking at the average loan amount would help to make the comparisons more material.

We also facet these plots by sectors to breakdown our analysis at a sector level.

```{r country facet sector fill btype, echo=TRUE, message=FALSE, warning=FALSE, results='asis', fig.width=15, fig.height=12}

# Creating graph for country vs amount facet by sector fill = Borrower type

loan_country_pc <- bor_chr %>% mutate(amount_per_cap= tot_loans/borr_count)

#Create country vector
country <- c("Peru", "Uganda","Bolivia","Cambodia","Paraguay","Mexico","United States","Tajikistan","Tanzania","Azerbaijan")

# Filter data for top 10 countries
loan_country <- subset(loan_country_pc, loan_country_pc$location.country %in% country)
loan_country  <- na.omit(loan_country)

#Bar Chart country vs Loan Amount
ggplot(loan_country) +
  aes(location.country, amount_per_cap, fill =bor_type) + 
  geom_bar(stat="identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 35, margin=margin(t= 10, r = 20, b = 0, l = 0))) +
  xlab("Country") +
  ylab("Average loan amount in USD") +
  scale_fill_OkabeIto(order = c(3,1,5)) +
  theme(axis.text.x = element_text(size=13), 
        axis.text.y = element_text(size=12),
        plot.title = element_text(size = 20),
        axis.title = element_text(size=16, face="bold")) +
  ggtitle("Average loan amount vs country by borrower type") +
  facet_wrap(~sector, ncol = 3) +
  theme(strip.text.x = element_text(size = 16)) + 
  labs(fill = "Borrower Type")

```

From the plot we can observe that average loan amount received by borrowers in the United States is much higher than other countries. However, due to this we are unable to gauge any trends for the other countries due to the variation in the scale. We try to remove United States from our filter and visualize this plot again.

```{r country facet sector w/o US, echo=TRUE, message=FALSE, warning=FALSE, results='asis', fig.width=15, fig.height=12}

# Creating graph for country vs amount facet by sector fill = Borrower type excluding United States

#Create country vector
country <- c("Peru", "Uganda","Bolivia","Cambodia","Paraguay","Mexico","Tajikistan","Tanzania","Azerbaijan")

# Filter data for top 10 countries
loan_country_woUS <- subset(loan_country_pc,loan_country_pc$location.country %in% country)
loan_country_woUS  <- na.omit(loan_country_woUS)

#Bar Chart country vs Loan Amount
ggplot(loan_country_woUS) +
  aes(location.country, amount_per_cap, fill = bor_type) + 
  geom_bar(stat="identity", position = "dodge")+
  theme(axis.text.x = element_text(angle = 35, margin=margin(t= 10, r = 20, b = 0, l = 0))) +
  xlab("Country") +
  ylab("Average loan amount in USD") +
  scale_fill_OkabeIto(order = c(3,1,5)) +
  theme(axis.text.x=element_text(size=13),
        axis.text.y=element_text(size=12),
        plot.title = element_text(size = 20),
        axis.title=element_text(size=16,face="bold")) +
  ggtitle("Average loan amount vs country by borrower type ") +
  facet_wrap(~sector, ncol = 3) +
  theme(strip.text.x = element_text(size = 16)) + 
  labs(fill="Borrower Type")

```

From this plot we can see more trends for the other countries. We can observe that loans in Personal Use and Entertainment Sector have received relatively lower average loan amounts for all borrower types across all countries.

------------------------------------------------------------------------

Next we look at how the average loan amount varies across countries within each sector with respect to the funding status of the loans.

```{r country facet sector fill fund stat, echo=TRUE, message=FALSE, warning=FALSE, results='asis', fig.width=15, fig.height=12}

#Bar Chart country vs Loan Amount
ggplot(loan_country) + 
  aes(location.country, amount_per_cap, fill = funding_status) +  
  geom_bar(stat="identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 35, margin=margin(t= 10, r = 20, b = 0, l = 0))) +
  xlab("Country") +
  ylab("Average loan amount in USD") +
  scale_fill_OkabeIto(order = c(2,6)) +
  theme(axis.text.x=element_text(size = 13),
        axis.text.y=element_text(size = 12),
        plot.title = element_text(size = 25),
        axis.title=element_text(size = 16,face = "bold")) +
  ggtitle("Average loan amount vs country by funding status") +
  facet_wrap(~sector, scales="free_y", ncol = 3) +
  theme(strip.text.x = element_text(size = 16)) + 
  labs(fill="Funding Status")

```

As we have already seen our data contains more funded loans than not funded loans, as is visible from our plot. From this plot we can see that all loans in the Education and Health sector have been funded. Another interesting observation is that in the Wholesale sector, United states has more Not funded loans than funded loans.

We can also look at how the funding status varies with respect to borrower type. We look at this at the sector level.

```{r bor_type facet fill funding status, echo = TRUE, warning = FALSE, message = FALSE, results='asis',fig.width=15, fig.height=12}

# Creating graph for bor_type vs amount facet by sector fill = Funding status

ggplot(loan_country) +
  aes(bor_type, amount_per_cap, fill = funding_status) +
  geom_bar(stat="identity", position = "dodge") +
  xlab("Borrower type") +
  ylab("Average loan amount in USD") +
  scale_fill_OkabeIto(order = c(2,6)) +
  theme(axis.text.x = element_text(size = 13),
        plot.title = element_text(size = 25),
        axis.title = element_text(size = 16, face = "bold")) +
  ggtitle("Average loan mount vs borrower type by funding status") +
  facet_wrap(~sector, ncol = 3) +
  theme(strip.text.x = element_text(size = 16)) + 
  labs(fill = "Funding Status")

```

From this visualization we observe that Female borrowers in the Arts, Education, Health, Retail and Services sector have higher funded loan amounts than the Male and Group borrowers. Group borrowers seem to have the lowest amount of not funded loans across all sectors. Male borrowers have a much higher amount of not funded loans than other borrower types.

We have looked at the borrower characteristics visually through our graphs and descriptive statistics. We now use Hypothesis Testing to further analyze our data.

### Hypothesis Testing

Hypothesis testing is a statistical inference method used to test an assumption about the data or more specifically the population parameters. We are interested in understanding how individual loans differ based on the gender of the applicant. The total loan amount is biased by the number of borrowers i.e. the number of female and group borrowers is much larger than the male borrowers, therefore we look at the average loan amount per borrower. To keep the analysis simple we subset our data to only Male and Female borrowers and test our hypothesis using a two-sample t-test. A t-test is a statistical test that compares the means of two samples. By subsetting and calculating the average loan amounts, we are comparing sample means using t-test to check our hypothesis.

Our null and alternative hypothesis are:

$H_{0} :$ $\texttt{The true difference between the average loan amount for Males and average loan amount for Females is equal to 0}$

$H_{1}:$ $\texttt{The true difference between the average loan amount for Males and average loan amount for Females is not equal to 0}$

-   $H_{0}$ is known as the null hypothesis and $H_{1}$ is know as the alternative hypothesis.\*

Before we proceed with the t-test, we need to check some assumptions to ensure we are performing a valid t-test. The assumptions for t-test are:

-   The data values are independent. The amount of the loan borrowed by one borrower does not depend on the loan amount borrowed by another.
-   We assume that number of borrowers in our data represent a random sample from the population of borrowers at Kiva.com
-   The values are continuous. The loan amount is a continuous variable
-   We assume that the data is normally distributed
-   We assume that the variance for men and women is equal

We will also check the normality and variance assumption before proceeding with the test.

```{r filter Male Female, out.width = '50%',echo = TRUE, fig.show='hold', warning = FALSE, message=TRUE}

# Subset data to Male and Female borrowers only
pc_gender <- subset(loan_country_pc, loan_country_pc$bor_type %in% c("Male","Female"))

par(mfrow=c(1,1)) 

#Plot distribution of average loan amount
ggplot(pc_gender, aes(x = amount_per_cap, color = bor_type, fill =bor_type)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 40) + 
  theme(legend.position = "none") +
  guides(color = FALSE) +
  ggtitle("Distribution of avg. loan amount by Gender") +
  xlab("Avg. Loan Amount in USD") +
  ylab("Count")

#Plot distribution of transformed average loan amount variable using log
ggplot(pc_gender, aes(x = log(amount_per_cap), color = bor_type, fill = bor_type)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 40) + 
  labs(fill = "Borrower Type") +
  guides(color = FALSE) +
  ggtitle("Distribution of log avg. loan amount by Gender") +
  xlab("Log of Avg. Loan Amount in USD") +
  ylab("Count")


```

From the histograms we can see that the loan amount is highly right skewed. We transform our loan amount variable using a log transformation and we can observe that the distribution of the transformed variable is more normally distributed. This the case for both Male and Female borrower type.

We check the variance assumption by calculating the standard deviation and variance by using the `sd` and `var` function on the transformed data.

```{r summary statistics,echo=TRUE,results='asis',warning=FALSE,message=FALSE}
sum_stat <- pc_gender %>% 
  group_by(bor_type) %>% 
  summarize (mean = round(mean(amount_per_cap),2),
             mean_log = round(mean(log(amount_per_cap)),2),
             variance = round(var(amount_per_cap),2),
             variance_log = round(var(log(amount_per_cap)),2),
             std_dev = round(sd(amount_per_cap),2),
             std_dev_log = round(sd(log(amount_per_cap)),2))

sum_stat %>% kbl(caption = "Borrower type summary statistics") %>% 
  kable_classic(bootstrap_options ="striped",full_width = F, html_font = "helvetica",position = "left")

```

We can see that the variance values and standard deviation are not equal or similar. However, we do see that they are similar for the transformed variable. Since the variances are very different, our assumption is not satisfied. However we can still test our hypothesis using Welch's T-test with unequal variances.

```{r, include = FALSE}
var.test(amount_per_cap ~ bor_type, data = pc_gender)$p.value
```

To perform the t-test we first separate our data into the respective groups of female and male borrowers. We use the `t-test` function to perform the t-test and calculate the t-statistic and p-values.

```{r hypothesis test, echo = TRUE, message=FALSE,warning=FALSE}

#Subset df to female borrowers
female <- subset(pc_gender,pc_gender$bor_type == "Female")
#Create df with only loan amount per borrower
female_loan <- female %>% pull(amount_per_cap)

#Subset df to male borrowers
male <- subset(pc_gender,pc_gender$bor_type == "Male")
#Create df with only loan amount per borrower
male_loan <- male %>% pull(amount_per_cap)

#hypothesis test
hyp <- t.test(female_loan,male_loan)
hyp
```

Our p-value is less than the alpha value 0.05. This implies that it is statistically significant. A statistically significant p-value indicates strong evidence against our null hypothesis. Therefore, we reject the null hypothesis that there is no difference in the sample means of average loan amount of male and female borrowers.

## Limitations and Conclusion

Our analysis has a few potential limitations. We have made a few assumptions when creating our indicator variables borrower type and funding status. We have also excluded some values in the status column while creating funding status. It is possible that those observations could have provided us with some more insight. Another limitation is the size of the observations for Not funded loans. There are very few Not Funded loans when compared to Funded Loans and this hinders any comparison we would like to make. We need to assume that our data does not contain the entire population of Not Funded loans.

To conclude, through descriptive statistics we observed that Food and Retail sectors have the highest number of funded loans. We also looked at the borrower types across sectors among the top 10 lending countries during 2005-2012. We noted that borrowers in the United States tend to receive a higher average loan amount than other countries. Through our hypothesis test we learned that the average loan amount received by a borrower does differ based on the borrower's gender.

Performing a paired t-test is one way of testing our hypothesis, however there are multiple other methods that can be explored to test this hypothesis and further analyze this data.
