---
title: "Survival Analysis"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(digits = 3)
```

Another way of approaching the question of what characteristics impact loan repayment is to utilize survival analysis, which permits us to retain more of the data rather than restricting our analysis to completed loans. We are also interested in how the status of loan tends to change over time. In this tutorial, we use the $\texttt{survival}$ package for survival analysis, which focuses on the expected duration of time until occurrence of an event of interest (in our case the repayment of loan).

IT WOULD MAKE A LOT MORE SENSE TO MODEL TIME TO DEFAULT, SO THAT SURVIVORS ARE LOANS THAT NEVER DEFAULT. YOU NEED TO EXPLAIN CENSORING IN THIS PARAGRAPH.

# Set Up

We load the packages, read in the data set, and manipulated the data as we did in the frequentist logistic regression section [**here**](link%20to%20be%20added%20after%20embedded%20in%20the%20website)

YOU DO NOT DO THE SAME MANIPULATION AS IN THE FREQUENTIST SECTION. IN THE FREQUENTIST SECTION YOU RESTRICTED IT TO COMPLETED LOANS WHICH DOES NOT HAVE ANY CENSORING SO YOU DIDN'T NEED SURVIVAL ANALYSIS.

```{r}
loans <- read.csv("data/loans.csv")
```

```{r include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(countrycode)
library(survival)
library(survminer)
```

# Survival Analysis

## Calculating Survival Times

We computed a new variable `span`, which is the time between when the loan is funded and when the loan is paid back, and stored it in a new data set `ds`.

```{r}
ds <- loans %>% 
  filter(status == "paid"|status == "defaulted") %>% 
  mutate(paid_amount = if_else(is.na(paid_amount), 0, paid_amount),
         continent = countrycode(sourcevar = location.country,
                            origin = "country.name",
                            destination = "continent"),
         span = (paid_yr - funded_yr) * 365 + (paid_mo - funded_mo) * 30 + (paid_day-funded_day),
         sex = case_when(borrower_m_count == 0 ~ "female",
                         borrower_f_count == 0 ~ "male",
                         TRUE ~ "mixed"),
         status = if_else(status == "paid", 1, 0)) %>% 
  select(status, span, sector, continent, sex)
```

## Fitting Survival Curves

The `surv()` function takes the variables `span` and `status` as input and creates an object which serves as the input of `survfir()` function. We first created a simple survival curve that doesn't consider any different groupings, by specifying just an intercept (\~1) in the formula in `survfit()` function.

```{r}
surv <- survfit(Surv(span,status==1)~1, data = ds)

ggsurvplot(
  fit = survfit(Surv(span,status==1)~1, data = ds),
  xlab="Number of days before the loan is repaid", 
  ylab="Overall survival probability")
```

Here, the x-axis specifies the number of days before the loan is repaid and the y-axis specifies the probability of loan being not repaid.

## Estimating Survival Probabilities

One quantity of interest in a survival analysis is the probability of surviving beyond a certain number of years. For example, to estimate the probability of loan not being repaid to 1 year, we can use `summary()` with the `time` argument. The `span` variable in our data is in days, so we are using `time = 365n`:

```{r}
summary(surv, time = 365/4)
summary(surv, time = 365/2)
summary(surv, time = 365)
summary(surv, time = 365*2)
```

As the output shows, the probability of borrowers not paying back the loan to a quarter year is 98%, to half a year is 69.6%, to 1 year is 26.4%, and to 2 years is 0.554%. Thus, we can conclude that almost all borrowers don't pay back the loan in the first quarter after they receive the funds; the majority of borrowers would repay the loan in 1 year; after 2 years only few borrowers haven't paid back their loan.

## Comparing Survival Times between Groups

We can take a look at the survival curve by loan sectors, continent, and sex of borrowers respectively:

```{r}
ggsurvplot(
  fit = survfit(Surv(span,status==1)~sector, data = ds),
  legend.title = "Sector",
  legend.labs = sort(unique(ds$sector)),
  title = "Survival Curves by Sector",
  xlab = "Number of days before the loan is repaid", 
  ylab = "Overall survival probability")

ggsurvplot(
  fit = survfit(Surv(span,status==1)~continent, data = ds),
  legend.title = "Continent",
  legend.labs = sort(unique(ds$continent)),
  title = "Survival Curves by Continent",
  xlab="Number of days before the loan is repaid", 
  ylab="Overall survival probability")

ggsurvplot(
  fit = survfit(Surv(span,status==1)~sex, data = ds),
  legend.title = "Sex",
  legend.labs = sort(unique(ds$sex)),
  title = "Survival Curves by Sex",
  xlab="Number of days before the loan is repaid", 
  ylab="Overall survival probability")
```

There seems to be no big difference in time of repayment across sectors. As for the continent, borrowers in Oceania tend to complete their loan repayments about a year later than borrowers in other continents. A borrower group that consists of both males and females tends to complete the loan repayment in the shortest period, and a female borrower or a female-only borrower group generally takes shorter time period to complete the loan repayment than a male borrower or a male-only borrower group does.

For numerical comparisons, we can use the `summary()` function to display the mean and median time it takes for each group to complete the loan repayment. For example, for `sex`:

```{r}
summary(survfit(Surv(span,status==1)~sex, data = ds))$table
```

The median time is 337 days for a male borrower or a male-only borrower group, 251 days for a female borrower or a female-only borrower group to fully repay the loan, and 165 days for a mixed-gender borrower group. There appears to be a shorter period to complete the loan repayment for mixed-gender borrower groups compared to single-gender borrower groups, and female borrowers tend to fully pay back their loans in a shorter time than male borrowers. This conclusion is consistent with our logistic model under the frequentist approach ([**here**](link%20to%20be%20added%20after%20embedded%20in%20the%20website)).

SUMMARIZE YOUR LOGISTIC REGRESSION FINDINGS AGAIN IN CASE THEY FORGOT.

YOU'RE DISCUSSING REGRESSION RESULTS LIKE DESCRIPTIVE STATISTICS. YOU COULD GIVE MEDIAN TIMES WITHOUT RUNNING ANY KIND OF MODEL. TRY TO FIND A TUTORIAL ON SURVIVAL ANALYSIS TO GET AN IDEA OF HOW TO INTREPRET COEFFICIENTS.
